Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-28 paper/paper.tex
fix on table

\begin{table*}[t!]
  \small
  \caption{\label{table:attacks}List of attacks.}
  \centering
  \begin{tabular}{llllp{10cm}}
    \toprule
    \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{Source} &
    \multicolumn{1}{c}{Type} &
    \multicolumn{1}{c}{Protocol} &
    \multicolumn{1}{c}{Groud Truth (relevant/checked options)} \\
    \midrule     
    Ping Scan & NMAP & AREC & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
    SYN Flood & hping3 & DoS & TCP & \textbf{flags}: S,12;
    \textbf{threshold}: type both, track by\_dst, count 5000, seconds
    5;\\
    UDP Teardrop & \Fix{WireShark} & DoS & UDP & \textbf{fragbits}:M; \textbf{id}:124; \\
    UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds
    5; \\
    Black Nurse & NetReseC & DoS & ICMP & \textbf{icode}:3; \textbf{itype}:3; \\
    \midrule
    JBoss JMX\MyComment{ Console Beanshell Exploit} & Nikto2 & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
    WeBid \MyComment{Local File Inclusion} & RFI & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
    ColdFusion \MyComment{PESC} & Nikto2 & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
    .htaccess access\MyComment{attempted-recon} & Nikto2 & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
    .idq access\MyComment{EXPLOIT ISAPI} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
    iisadmin access\MyComment{web-application-attack} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & Nikto2 & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
    Apache Struts 2 & WRCCDC & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
    WordPress Login & WRCCDC & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

    \midrule
    Ping Scan & 6 & 5 & 2 & - \\
    SYN Flood & 4 & 2 & 2 & - \\
    JBoss JMX & 5 & 23 & 228 & \Fix{23} \\
    WeBid & 1 & 23 & - & -\\    
    ColdFusion & 7 & 19 & 34 & \Fix{5}\\
    .htaccess access & 5 & 17 & 984 & \Fix{2}\\
    .idq access & 14 & 19 & 511 & \Fix{2} \\
    iisadmin access & 4 & 17 & \Fix{861} & \Fix{1} \\
    /system32/ in URI & 4 & 20 & 1,913 & \Fix{2} \\
    Script tag in URI & 23 & 22 & 786 & - \\
    Apache Struts 2 & - & - & - & - \\
    Wordpress Login & - & - & - & - \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~Recall that \tname\ takes
pcap files on input. As explained on Section~\ref{sec:minimization},
these files abstract network messages~\cite{pcap} and can be used to
avoid network traffic during rule synthesis. For some of the attacks
listed on Table~\ref{table:attacks}, we obtained pcap files directly
from trusted security websites (\eg{},
NetReseC~\cite{pcap-attacks}\Fix{confirm}).\Mar{reminder: we still
  don't have any example of attack listed on Table~\ref{table:attacks}
  whose pcap file was obtained from those websites.} \Luc{Ataques Apache 
  e WordPress. @Guilherme adicionar fontes} \Gui{O NetReseC oferece uma lista de sites que oferecem datasets, como o proprio Stratosphere, que estamos utilizando para trafego normal. Para esses dois ataques novos com content, peguei do WRCCDC - Western Regional Collegiate Cyber Defense Competition: https://archive.wrccdc.org/pcaps/}In addition to
these sources, we used the Nikto2~\cite{nikto} open source security
tool as follows to obtain pcap files for (non-flooding) http and tcp
attacks. We configured an http server (Apache2) and ran Nikto2 to
attack that server. Simultenously, we ran
WireShark~\cite{wireshark-net-monitor} to record the network traffic
and Suricata to analyze the traffic. We configured \suri\ with all
rules from the ``Emerging Threats Open
Ruleset''~\cite{emerging-threats-open} and configured Nikto2 with
default parameters, which makes it replay all recorded attacks. After
execution, we inspected the \suri\ log to identify which rules have
been activated and which packets were involved on each
activation. With the packet numbers, it is possible to recover the
corresponding pcap files, using Wireshark, as to ran \tname{} on
them. It is worth noting that some rules are triggered multiple times,
indicating that Nikto2 expresses multiple variants of the same
attack. RQ3 evaluates the importance of these variants to \tname{}'s
performance. We used similar methodology to obtain pcap files using
other security testing tools. We used NMAP~\cite{netmap} to reproduce
the Ping Scan attack (see Section~\ref{sec:active-recon}) and
hping3~\cite{hping3} to reproduce the SYN Flood attack (see
Section~\ref{sec:dos}).

>>>>

\begin{table*}[t!]
  \small
  \caption{\label{table:attacks}List of attacks.}
  \centering
  \begin{tabular}{llllp{10cm}}
    \toprule
    \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{Source} &
    \multicolumn{1}{c}{Type} &
    \multicolumn{1}{c}{Protocol} &
    \multicolumn{1}{c}{Groud Truth (relevant/checked options)} \\
    \midrule     
    Ping Scan & NMAP & AREC & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
    SYN Flood & hping3 & DoS & TCP & \textbf{flags}: S,12;
    \textbf{threshold}: type both, track by\_dst, count 5000, seconds
    5;\\
    UDP Teardrop & \Fix{WireShark} & DoS & UDP & \textbf{fragbits}:M; \textbf{id}:124; \\
    UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds
    5; \\
    Black Nurse & NetReseC & DoS & ICMP & \textbf{icode}:3; \textbf{itype}:3;\textbf{threshold}: type both, track by\_dst, count 250, seconds
    1; \\
    \midrule
    JBoss JMX\MyComment{ Console Beanshell Exploit} & Nikto2 & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
    WeBid \MyComment{Local File Inclusion} & RFI & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
    ColdFusion \MyComment{PESC} & Nikto2 & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
    .htaccess access\MyComment{attempted-recon} & Nikto2 & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
    .idq access\MyComment{EXPLOIT ISAPI} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
    iisadmin access\MyComment{web-application-attack} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & Nikto2 & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
    Apache Struts 2 & WRCCDC & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
    WordPress Login & WRCCDC & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

    \midrule
    Ping Scan & 6 & 5 & 2 & - \\
    SYN Flood & 4 & 2 & 2 & - \\
    UDP Teardrop & - & - & - & - \\
    UDP Flood & - & - & - & - \\
    Black Nurse & - & - & - & - \\
    JBoss JMX & 5 & 23 & 228 & \Fix{23} \\
    WeBid & 1 & 23 & - & -\\    
    ColdFusion & 7 & 19 & 34 & \Fix{5}\\
    .htaccess access & 5 & 17 & 984 & \Fix{2}\\
    .idq access & 14 & 19 & 511 & \Fix{2} \\
    iisadmin access & 4 & 17 & \Fix{861} & \Fix{1} \\
    /system32/ in URI & 4 & 20 & 1,913 & \Fix{2} \\
    Script tag in URI & 23 & 22 & 786 & - \\
    Apache Struts 2 & - & - & - & - \\
    Wordpress Login & 8 & 59 & - & - \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~Recall that \tname\ takes
pcap files on input. As explained on Section~\ref{sec:minimization},
these files abstract network messages~\cite{pcap} and can be used to
avoid network traffic during rule synthesis. For some of the attacks
listed on Table~\ref{table:attacks}, we obtained pcap files directly
from trusted security websites (\eg{},
NetReseC~\cite{pcap-attacks}\Fix{confirm}).\Mar{reminder: we still
  don't have any example of attack listed on Table~\ref{table:attacks}
  whose pcap file was obtained from those websites.} \Luc{Ataques Apache 
  e WordPress. @Guilherme adicionar fontes} \Gui{O NetReseC possui uma lista de sites (alem dos pcaps que o proprio site ja oferece, como foi obtido o Black Nurse) que oferecem datasets, como o proprio Stratosphere, que estamos utilizando para trafego normal. Para esses dois ataques novos com content, peguei do WRCCDC - Western Regional Collegiate Cyber Defense Competition: https://archive.wrccdc.org/pcaps/}In addition to
these sources, we used the Nikto2~\cite{nikto} open source security
tool as follows to obtain pcap files for (non-flooding) http and tcp
attacks. We configured an http server (Apache2) and ran Nikto2 to
attack that server. Simultenously, we ran
WireShark~\cite{wireshark-net-monitor} to record the network traffic
and Suricata to analyze the traffic. We configured \suri\ with all
rules from the ``Emerging Threats Open
Ruleset''~\cite{emerging-threats-open} and configured Nikto2 with
default parameters, which makes it replay all recorded attacks. After
execution, we inspected the \suri\ log to identify which rules have
been activated and which packets were involved on each
activation. With the packet numbers, it is possible to recover the
corresponding pcap files, using Wireshark, as to ran \tname{} on
them. It is worth noting that some rules are triggered multiple times,
indicating that Nikto2 expresses multiple variants of the same
attack. RQ3 evaluates the importance of these variants to \tname{}'s
performance. We used similar methodology to obtain pcap files using
other security testing tools. We used NMAP~\cite{netmap} to reproduce
the Ping Scan attack (see Section~\ref{sec:active-recon}) and
hping3~\cite{hping3} to reproduce the SYN Flood attack (see
Section~\ref{sec:dos}).
